01_LH-SIP_Data_Reduction

### 2023-01-06 01_LH-SIP_Data_Reduction.Rmd
select relevant columns for export

## Load sourced functions
These functions are found in the `libs/` directory.
```{r}
# Sourced Functions
source(file.path("libs", "visualization.R"))              # Visualization scripts
source(file.path("libs", "chromeleon.R"))                 # Chromeleon Reader
source(file.path("libs", "error_prop.R"))                 # Error Propagation
source(file.path("libs", "calculate_turnover.R"))         # Turnover Calculator
source(file.path("libs", "d2H_to_2F.R"))                  # Convert delta to frac abund
source(file.path("libs", "rename_FAs.R"))                 # Rename FAs to correctly ID'd FAs
`%nin%` = Negate(`%in%`)                                  # Nifty "Not In" function
var_to_str = function(v) {return(deparse(substitute(v)))} # Convert variable name to string
```

@@

# EXCLUDE RUN WE'VE DEEMED PROBLEMATIC
sample_files <- iso_filter_files(sample_files, analysis %nin% problematic_analyses_num)

@@

## Load memory-corrected IRMS data

```{r}
LH_SIP_memcorr <- readxl::read_excel("data/samples_data_memory_corrected.xlsx") %>% 
# Trim it down
  select(-c(standard, is_std_peak, folder, file_datetime, id2, individual, injection_volume,
            gc_method, map_id, gc_ramp, group, peak_type, ref_nr, calib_peak, peak_nr,
            is_ref, rt_start, rt, rt_end, n_overlapping, n_matches, use_in_calib, d2H_in_calib, d2H_resid, 
            d2H_calib_points)) %>% 
  filter(id1 %nin% c("cr0f0x", "cr0f0y", "cr0f0z",
                     "cr0f3x", "cr0f3y", "cr0f3z",
                     "cr0f7x", "cr0f7y", "cr0f7z",
                     "cri0", "cri3", "cri7")) %>% 
  rename(sample_id = id1,
         Analysis = analysis) %>% 
  select(Analysis, sample_id, compound, area2, area3, note, 
          calibrated_d2H_without_memory, 
          calibrated_d2H_without_memory_se,
          calibrated_d2H_with_memory_correction, 
          calibrated_d2H_with_memory_correction_se) %>% 
  mutate(Analysis = as.numeric(Analysis)) %>% 
  mutate(
    calibrated_at2H_without_memory =
      d2H_to_2F(calibrated_d2H_without_memory),
    calibrated_at2H_without_memory_se =
      d2H_to_2F(calibrated_d2H_without_memory_se),
    calibrated_at2H_with_memory_correction =
      d2H_to_2F(calibrated_d2H_with_memory_correction), 
    calibrated_at2H_with_memory_correction_se =
      d2H_to_2F(calibrated_d2H_with_memory_correction_se)
  )
```

@@

## Assimilation efficiency calcs

@@

FID_results_abundance_no_stds <- FID_results_abundance %>% 
  filter(
    compound %nin% c("21:0 (STD)", "23:0 (STD)", "PAIBE"),
    str_detect(sample_id, "FAME", negate = TRUE),
    str_detect(sample_id, "BAME", negate = TRUE)
  )
```

@@

# Filter out runs that are on our problematic_runs list
peaks_mapped <- peaks_mapped %>% 
  filter(Analysis %nin% problematic_runs)
```

@@

```{r}
# Reduce the 0d --> 3d --> 7d incubation data
peaks_mapped_with_zeros_analytes_reduced <- peaks_mapped_with_zeros_analytes %>%
  # Add assimilation efficiency data
  mutate(
    a = d2H_assim_lm_summary$a_mn,
    sa = d2H_assim_lm_summary$sa
  ) %>% 
  # Column reduction
  select(Analysis,
         sample_id,
         terrain,
         soil,
         inc_time_d,
         inc_time_d_str,
         compound,
         area_FID,
         rel_area_FID,
         rt,
         d2H_mn_zero,
         d2H,
         at2H_mn_zero,
         at2H_mn_zero_mc, # MC
         at2H, # not-corrected value
         calibrated_at2H_without_memory,
         calibrated_at2H_with_memory_correction, # MC
         area2,
         area3,
         area_FID,
         compound_ug_per_g_soil,
         a, # assimilation efficiency
         `2F_Me`,
         `2F_FAME`,
         `2F_FAME_mc`, # MC
         `2F_alk`,
         `2F_alk_mc`,  # MC
         F_label_mn,
         # Uncertainty terms:
         F_label_se, # error in the label strength (pipetting error)
         F_0_at2H_se, # error in the initial F_0 (instrument error)
         calibrated_at2H_without_memory_se, # error in F_T
         calibrated_at2H_with_memory_correction_se, # error in F_T, MC
         sa, # error in assimilation efficiency (se of literature values)
  ) %>% 
  # Turn all unit dbls into unitless
  mutate(
    across(
      where(is.numeric),
      as.numeric
    )
  ) %>% 
  # Rename FAs to match correctly ID'd analytes
  rename_FAs()

>>>>

## Load sourced functions
These functions are found in the `libs/` directory.
```{r}
# Sourced Functions
source(file.path("libs", "visualization.R"))              # Visualization scripts
source(file.path("libs", "chromeleon.R"))                 # Chromeleon Reader
source(file.path("libs", "error_prop.R"))                 # Error Propagation
source(file.path("libs", "calculate_turnover.R"))         # Turnover Calculator
source(file.path("libs", "d2H_to_2F.R"))                  # Convert delta to frac abund
source(file.path("libs", "rename_FAs.R"))                 # Rename FAs to correctly ID'd FAs
var_to_str = function(v) {return(deparse(substitute(v)))} # Convert variable name to string
```

@@

# EXCLUDE RUN WE'VE DEEMED PROBLEMATIC
sample_files <- iso_filter_files(sample_files, !analysis %in% problematic_analyses_num)

@@

## Load memory-corrected IRMS data!

```{r}
LH_SIP_memcorr <- readxl::read_excel("data/samples_data_memory_corrected_20221228.xlsx") %>% 
# Trim it down
  select(-c(type, # or `standard` if using original mc dataset 
            #is_std_peak,
            folder, 
            file_datetime, 
            id2, 
            #individual,
            #injection_volume,
            #gc_method,
            #map_id, 
            gc_ramp, 
            group, 
            #peak_type,
            #ref_nr,
            #calib_peak, 
            peak_nr,
            #is_ref,
            rt_start,
            rt, 
            rt_end, 
            n_overlapping,
            n_matches,
            #use_in_calib,
            #d2H_in_calib, 
            #d2H_resid, 
            d2H_calib_points
            )
         ) %>% 
  filter(!id1 %in% c("cr0f0x", "cr0f0y", "cr0f0z",
                     "cr0f3x", "cr0f3y", "cr0f3z",
                     "cr0f7x", "cr0f7y", "cr0f7z",
                     "cri0", "cri3", "cri7")) %>% 
  rename(sample_id = id1,
         Analysis = analysis) %>% 
  select(Analysis, sample_id, compound, area2, area3, note,
          below_area_range, 
          calibrated_d2H_without_area,
          calibrated_d2H_without_area_se,
          calibrated_d2H_without_memory, 
          calibrated_d2H_without_memory_se,
          calibrated_d2H_with_memory_correction, 
          calibrated_d2H_with_memory_correction_se) %>% 
  mutate(Analysis = as.numeric(Analysis)) %>% 
  # Convert Delta Values in Permil (vs. VSMOW) to fractional abundance (at%)
  mutate(
    # Without area correction:
    calibrated_at2H_without_area = d2H_to_2F(calibrated_d2H_without_area),
    calibrated_at2H_without_area_se = d2H_to_2F(calibrated_d2H_without_area_se),
    # With Area, without memory correction:
    calibrated_at2H_without_memory = d2H_to_2F(calibrated_d2H_without_memory),
    calibrated_at2H_without_memory_se = d2H_to_2F(calibrated_d2H_without_memory_se),
    # With memory correction:
    calibrated_at2H_with_memory_correction = d2H_to_2F(calibrated_d2H_with_memory_correction), 
    calibrated_at2H_with_memory_correction_se = d2H_to_2F(calibrated_d2H_with_memory_correction_se)
  )
```

@@

@@

FID_results_abundance_no_stds <- FID_results_abundance %>% 
  filter(
    !compound %in% c("21:0 (STD)", "23:0 (STD)", "PAIBE"),
    str_detect(sample_id, "FAME", negate = TRUE),
    str_detect(sample_id, "BAME", negate = TRUE)
  )
```

@@

# Filter out runs that are on our problematic_runs list
peaks_mapped <- peaks_mapped %>% 
  filter(!Analysis %in% problematic_runs)
```

@@

```{r}
# Reduce the 0d --> 3d --> 7d incubation data
peaks_mapped_with_zeros_analytes_reduced <- peaks_mapped_with_zeros_analytes %>%
  # Add assimilation efficiency data
  mutate(
    a = d2H_assim_lm_summary$a_mn,
    sa = d2H_assim_lm_summary$sa
  ) %>% 
  # REPAIR out-of-area t=0 datapoints:
  mutate(
    calibrated_at2H_with_memory_correction =
      case_when(
        below_area_range == TRUE & inc_time_d == 0 ~ at2H_mn_zero,
        TRUE ~ calibrated_at2H_with_memory_correction
        ),
    below_area_range = 
      case_when(
        below_area_range == TRUE & inc_time_d == 0 ~ FALSE,
        below_area_range == TRUE ~ TRUE,
        below_area_range == FALSE ~ FALSE
      )
  ) %>% 
  # REMOVE peaks that are outside area range of accurate correction!
  filter(below_area_range == FALSE) %>% 
  # Column reduction
  select(Analysis,
         sample_id,
         terrain,
         soil,
         inc_time_d,
         inc_time_d_str,
         compound,
         area_FID,
         rel_area_FID,
         rt,
         d2H_mn_zero,
         d2H,
         at2H_mn_zero,
         at2H_mn_zero_mc, # MC
         at2H, # not-corrected value
         calibrated_at2H_without_area, # F_T, no area corr
         calibrated_at2H_without_memory, # F_T, no memcorr
         calibrated_at2H_with_memory_correction, # F_T, MC!
         area2,
         area3,
         area_FID,
         compound_ug_per_g_soil,
         a, # assimilation efficiency
         `2F_Me`,
         `2F_FAME`,
         `2F_FAME_mc`, # MC
         `2F_alk`,
         `2F_alk_mc`,  # MC
         F_label_mn,
         # Uncertainty terms:
         F_label_se, # error in the label strength (pipetting error)
         F_0_at2H_se, # error in the initial F_0 (instrument error)
         calibrated_at2H_without_area_se, # error in F_T, no area corr
         calibrated_at2H_without_memory_se, # error in F_T, no memcorr
         calibrated_at2H_with_memory_correction_se, # error in F_T, MC!
         sa, # error in assimilation efficiency (se of literature values)
  ) %>% 
  # Turn all unit dbls into unitless
  mutate(
    across(
      where(is.numeric),
      as.numeric
    )
  ) %>% 
  # Rename FAs to match correctly ID'd analytes
  rename_FAs()
